---
title: "R Notebook"
author: ruonanjia/EW
---


```{r}
rm(list = ls())
```

```{r}
# Download package tarball from CRAN archive
url <- "https://cran.r-project.org/src/contrib/Archive/zipcode/zipcode_1.0.tar.gz"
pkgFile <- "zipcode_1.0.tar.gz"
download.file(url = url, destfile = pkgFile)
# Install dependencies
install.packages(c("ada", "ipred", "evd","sjmisc"))
# Install package
install.packages(pkgs=pkgFile, type="source", repos=NULL)
# Delete package tarball
unlink(pkgFile)
```


load libraries
```{r}
library(ggplot2)
library(zipcode) # install only available through archive
library(maps)
library(ggpubr)
library("PerformanceAnalytics") # for chart correlation
library(usmap) # for plotting data us maps
library(scales) # for muted() function
library(tidyverse)
library(dplyr)
# library(memisc)
library(MASS)
```

# load data
```{r}
setwd("C:/Users/ekw28/ekw_lab_projects/dm_covid/data")

data.all <- read.csv("data_all_sub.csv")
tracker <- read.csv("all_states_excel_formatted.csv")
state_name <- read.csv("state_name.csv")
# state_pops <- read.csv("state_pops.csv")
data.all$age <- as.numeric(as.character(data.all$age)) + 17 # correct coding
#rm(data.attitude)
```

# Merge tracker and state_name
```{r}
# typeof(tracker)
# tracker
# data.all
# state_name
colnames(state_name)[2] <- "State"
tracker <- merge(tracker, state_name, by="State")
tracker
```

# Preprocess some data columns, run the following chunks before further analysis:

Separate young and old, compare with 60 yr
```{r}
data.all$is.young <- 1
data.all$is.young[data.all$age >= 60] = 0 
data.all$is.young <- as.factor(data.all$is.young)
```

# clean zipcode
```{r}
data.all$zip_postal_code <- as.numeric(as.character(data.all$zip_postal_code))
data(zipcode)
data.all$zipcode<- clean.zipcodes(data.all$zip_postal_code)
```

# state based on zip
```{r}
colnames(zipcode)
colnames(zipcode)[1] = "zipcode"
data.all <- merge(data.all, zipcode, by="zipcode")
```

# correct tracker week
```{r}
# in the original data sheet, week 1 refers to the most recent week
typeof(tracker$Week)
tracker$Week <- 7-tracker$Week
```

# sort data.all start dates into corresponding weeks
```{r}
Week <- c(1,2,3,4,5,6)
date <- c("2020-03-29", "2020-04-05", "2020-04-12", "2020-04-19", "2020-04-26", "2020-05-03")
timing <- data.frame(Week, date)
timing$Week <- as.integer(timing$Week)
tracker <- merge(tracker, timing, by="Week")
```

# change column names
```{r}
# colnames(tracker)[1] <- "Week"
colnames(tracker)[14] <- "state" 
  #lowercase "s" = abbr ([14]), uppercase "s" = full name ([2])
colnames(tracker)
```

# change data type
```{r}
tracker$state <- as.character(tracker$state)
tracker$Week <- as.integer(as.character(tracker$Week))
# tracker$testing <- as.numeric(tracker$testing)
# no data for AL
# typeof(tracker$testing)
```

End of preprocessing

# match date in data.all with week in tracker (?)
```{r}
# data_all_sub$Week <- NA
data.all$Week <- NA
for (i in 1:nrow(data.all)) { 
  if (data.all$state[i] != "") {
    data.all$Week[i] <-
      tracker$Week[as.Date(data.all$StartDate[i], "%m/%d/%Y")>=as.Date(tracker$date) &
                        as.Date(data.all$StartDate[i], "%m/%d/%Y")<as.Date(tracker$date)+7 &
                        as.character(data.all$state[i]) == tracker$state]
  }
}
```


# combine data with case tracker, matching time
```{r}
# data.all$testing <- NA
data.all$Avg.positive.tests <- NA
data.all$Avg.total.test.results <- NA
data.all$Avg.deaths <- NA
data.all$Avg.cumulative.hospitalized <- NA
data.all$Avg.current.hospitalized <- NA
data.all$Avg.hospitalized.increase <- NA
data.all$week <- NA

# if within the week
for (i in 1:nrow(data.all)) {
  # if (data.all$state[i] != "") {
    # data.all$testing[i] <-
    #   tracker$testing[as.Date(data.all$StartDate[i])>=as.Date(tracker$date) &
    #                     as.Date(data.all$StartDate[i])<as.Date(tracker$date)+7 &
    #                     as.character(data.all$state[i]) == tracker$state]
    
    data.all$Avg.positive.tests[i] <-
      tracker$Avg.positive.tests[as.Date(data.all$StartDate[i],"%m/%d/%Y")>=as.Date(tracker$date) &
                        as.Date(data.all$StartDate[i],"%m/%d/%Y")<as.Date(tracker$date)+7 &
                        as.character(data.all$state[i]) == tracker$state]
    # first line = starting week of suvey > tracker date
    # second line = starting date of survey < tracker date + 7
    # AKA identify time interval 
    # third line = state of survey = matched to the tracker state
    # all conditions are linked by logical operator "&"
    data.all$Avg.total.test.results[i] <-
      tracker$Avg.total.test.results[as.Date(data.all$StartDate[i],"%m/%d/%Y")>=as.Date(tracker$date) &
                        as.Date(data.all$StartDate[i],"%m/%d/%Y")<as.Date(tracker$date)+7 &
                        as.character(data.all$state[i]) == tracker$state]
    
    data.all$Avg.deaths[i] <-
      tracker$Avg.deaths[as.Date(data.all$StartDate[i],"%m/%d/%Y")>=as.Date(tracker$date) &
                        as.Date(data.all$StartDate[i],"%m/%d/%Y")<as.Date(tracker$date)+7 &
                        as.character(data.all$state[i]) == tracker$state]
    
    data.all$Avg.cumulative.hospitalized[i] <-
      tracker$Avg.cumulative.hospitalized[as.Date(data.all$StartDate[i],"%m/%d/%Y")>=as.Date(tracker$date) &
                        as.Date(data.all$StartDate[i],"%m/%d/%Y")<as.Date(tracker$date)+7 &
                        as.character(data.all$state[i]) == tracker$state]
    
    data.all$Avg.current.hospitalized[i] <-
      tracker$Avg.current.hospitalized[as.Date(data.all$StartDate[i],"%m/%d/%Y")>=as.Date(tracker$date) &
                        as.Date(data.all$StartDate[i],"%m/%d/%Y")<as.Date(tracker$date)+7 &
                        as.character(data.all$state[i]) == tracker$state]
    
    data.all$Avg.hospitalized.increase[i] <-
      tracker$Avg.hospitalized.increase[as.Date(data.all$StartDate[i],"%m/%d/%Y")>=as.Date(tracker$date) &
                        as.Date(data.all$StartDate[i],"%m/%d/%Y")<as.Date(tracker$date)+7 &
                        as.character(data.all$state[i]) == tracker$state]
    
    data.all$week[i] <-
      tracker$Week[as.Date(data.all$StartDate[i],"%m/%d/%Y")>=as.Date(tracker$date) &
                        as.Date(data.all$StartDate[i],"%m/%d/%Y")<as.Date(tracker$date)+7 &
                        as.character(data.all$state[i]) == tracker$state]
  }

```


# change data type
```{r}
data.all$ambig1_react_state <- as.integer(as.character(data.all$ambig1_react_state))
data.all$ambig1_gov_local <- as.integer(as.character(data.all$ambig1_gov_local))
data.all$anx_feelingnow <- as.integer(as.character(data.all$anx_feelingnow))
```


START ANALYSIS 

# preliminary visualization of covid&dm dataset  
```{r}
# box plot: ambig1_react_state v. age
names1 <- c("Completely overreacting", "Somewhat overreacting", "Appropriate", "Somewhat insufficient", "Completely insufficient") #1-5
data.all$ambig1_react_state <- as.factor(data.all$ambig1_react_state)
plot1 <- ggplot(data.all, aes(x=ambig1_react_state, y=age)) + geom_boxplot() +  ggtitle("ambig1_react_state v. age") +
  theme(plot.title = element_text(hjust = 0.5))                                             
plot1 + scale_x_discrete(labels=names1) + theme(axis.text.x=element_text(angle=60,hjust=1))+ labs(x="rating", y="age")

# box plot: ambig1_react_econ v. age
names1 <- c("Completely overreacting", "Somewhat overreacting", "Appropriate", "Somewhat insufficient", "Completely insufficient") #1-5
data.all$ambig1_react_econ <- as.factor(data.all$ambig1_react_econ)
plot1 <- ggplot(data.all, aes(x=ambig1_react_econ, y=age)) + geom_boxplot() +  ggtitle("ambig1_react_econ v. age") +
  theme(plot.title = element_text(hjust = 0.5))                                             
plot1 + scale_x_discrete(labels=names1) + theme(axis.text.x=element_text(angle=60,hjust=1))+ labs(x="rating", y="age")

# box plot: ambig1_react_nation v. age
names1 <- c("Completely overreacting", "Somewhat overreacting", "Appropriate", "Somewhat insufficient", "Completely insufficient") #1-5
data.all$ambig1_react_nation <- as.factor(data.all$ambig1_react_nation)
plot1 <- ggplot(data.all, aes(x=ambig1_react_nation, y=age)) + geom_boxplot() +  ggtitle("ambig1_react_nation v. age") +
  theme(plot.title = element_text(hjust = 0.5))                                             
plot1 + scale_x_discrete(labels=names1) + theme(axis.text.x=element_text(angle=60,hjust=1))+ labs(x="rating", y="age")

# box plot: ambig1_react_science v. age
names1 <- c("Completely overreacting", "Somewhat overreacting", "Appropriate", "Somewhat insufficient", "Completely insufficient") #1-5
data.all$ambig1_react_science <- as.factor(data.all$ambig1_react_science)
plot1 <- ggplot(data.all, aes(x=ambig1_react_science, y=age)) + geom_boxplot() +  ggtitle("ambig1_react_science v. age") +
  theme(plot.title = element_text(hjust = 0.5))                                             
plot1 + scale_x_discrete(labels=names1) + theme(axis.text.x=element_text(angle=60,hjust=1))+ labs(x="rating", y="age")

# box plot: ambig1_react_worker v. age
names1 <- c("Completely overreacting", "Somewhat overreacting", "Appropriate", "Somewhat insufficient", "Completely insufficient") #1-5
data.all$ambig1_react_worker <- as.factor(data.all$ambig1_react_worker)
plot1 <- ggplot(data.all, aes(x=ambig1_react_worker, y=age)) + geom_boxplot() +  ggtitle("ambig1_react_worker v. age") +
  theme(plot.title = element_text(hjust = 0.5))                                             
plot1 + scale_x_discrete(labels=names1) + theme(axis.text.x=element_text(angle=60,hjust=1))+ labs(x="rating", y="age")

# box plot: ambig1_react_you v. age
names1 <- c("Completely overreacting", "Somewhat overreacting", "Appropriate", "Somewhat insufficient", "Completely insufficient") #1-5
data.all$ambig1_react_you <- as.factor(data.all$ambig1_react_you)
plot1 <- ggplot(data.all, aes(x=ambig1_react_you, y=age)) + geom_boxplot() +  ggtitle("ambig1_react_you v. age") +
  theme(plot.title = element_text(hjust = 0.5))                                             
plot1 + scale_x_discrete(labels=names1) + theme(axis.text.x=element_text(angle=60,hjust=1)) + labs(x="rating", y="age")

# geom violin: ambig1_react_state v. political typology
names1 <- c("Completely overreacting", "Somewhat overreacting", "Appropriate", "Somewhat insufficient", "Completely insufficient") # 1-5 rxn ratings ambig1_react parameter
names2 <- c("Very liberal","Liberal", "Slightly liberal", "Moderate/middle-of-the-raod","Slightly conservative", "Conservative","Very conservative","Don't know/not political","Libertarian","Other")
data.all$ambig1_react_state <- as.numeric(data.all$ambig1_react_state)
data.all$political_typology <- as.factor(data.all$political_typology)
plot <- ggplot(data.all, aes(x=ambig1_react_state, y=political_typology)) + geom_violin()
plot + stat_summary(fun.y=median, geom="point") + scale_x_continuous(labels=names1) + scale_y_discrete(labels=names2) + theme(axis.text.x=element_text(angle=60,hjust=1)) + labs(x="rating", y="political typology")

#geom_point
data.all$ambig1_react_state <- as.numeric(data.all$ambig1_react_state)
data.all$political_typology <- as.factor(data.all$political_typology)
plot <- ggplot(data.all, aes(x=factor(ambig1_react_state), y=political_typology)) + geom_point()
plot + scale_x_discrete(labels=names1) + scale_y_discrete(labels=names2) + theme(axis.text.x=element_text(angle=60,hjust=1)) + labs(x="rating", y="political typology")

# ambig1_react_you v. political typology
data.all$ambig1_react_you <- as.numeric(data.all$ambig1_react_you)
data.all$political_typology <- as.factor(data.all$political_typology)
plot <- ggplot(data.all, aes(x=ambig1_react_you, y=political_typology)) + geom_violin()
plot + stat_summary(fun.y=median, geom="point") + scale_x_continuous(labels=names1) + scale_y_discrete(labels=names2) + theme(axis.text.x=element_text(angle=60,hjust=1)) + labs(x="rating", y="political typology")

# ambig1_react_science v. political typology
data.all$ambig1_react_science <- as.numeric(data.all$ambig1_react_science)
data.all$political_typology <- as.factor(data.all$political_typology)
plot <- ggplot(data.all, aes(x=ambig1_react_science, y=political_typology)) + geom_violin()
plot + stat_summary(fun.y=median, geom="point") + scale_x_continuous(labels=names1) + scale_y_discrete(labels=names2) + theme(axis.text.x=element_text(angle=60,hjust=1)) + labs(x="rating", y="political typology")

# ambig1_react_worker v. political typology
data.all$ambig1_react_worker <- as.numeric(data.all$ambig1_react_worker)
data.all$political_typology <- as.factor(data.all$political_typology)
plot <- ggplot(data.all, aes(x=ambig1_react_worker, y=political_typology)) + geom_violin()
plot + stat_summary(fun.y=median, geom="point") + scale_x_continuous(labels=names1) + scale_y_discrete(labels=names2) + theme(axis.text.x=element_text(angle=60,hjust=1)) + labs(x="rating", y="political typology")

# ambig1_react_econ v. political typology
data.all$ambig1_react_econ <- as.numeric(data.all$ambig1_react_econ)
data.all$political_typology <- as.factor(data.all$political_typology)
plot <- ggplot(data.all, aes(x=ambig1_react_econ, y=political_typology)) + geom_violin()
plot + stat_summary(fun.y=median, geom="point") + scale_x_continuous(labels=names1) + scale_y_discrete(labels=names2) + theme(axis.text.x=element_text(angle=60,hjust=1)) + labs(x="rating", y="political typology")
```

# preliminary visualization using external metrics
  How do people's perceptions of how states are reacting change based on deaths, hospitalizations, and testing?
  
```{r}
# for each of large workers
  # violin plot with jitter overlaid 
    # x axis = rating of category
        # ambig1_react_worker
          # ambig1_react_worker
          # ambig1_react_econ
          # ambig1_react_science
          # ambig1_react_you
          # ambig1_react_nation
    # y axis (prioritize by significance and data quality)
      # avg.positive.tests
      # avg.total.test.results
      # avg.deaths
      # avg.cumulative.hospitalized
      # avg.hospitalized.increase
    # color 
      # week

# chage data type
data.all$ambig1_react_nation <- ordered(data.all$ambig1_react_nation)

# plot

ggplot(data.all[data.all$state == "CA", ], aes(x=ambig1_react_nation, y=Avg.deaths, color = week)) +
  geom_violin() +
  geom_jitter(width = 0.1) + ggtitle("CA") +  theme(plot.title = element_text(hjust = 0.5))
```

# death rate v. ambig1_react_category
  - violin plot with jitter overlaid 
      death rate (total deaths/total state pop)
      ambig1_react_categories
  - calculate national death rate by week (deaths/total pop)
      national pop metric: https://www.census.gov/popclock/ (2019)
      
```{r}
# calculate national death rate for e/ week

index1 <- 1:50
index2 <- 51:100
index3 <- 101:150
index4 <- 151:200
index5 <- 201:250
index6 <- 251:300

sort(tracker$Week)

# sum avg deaths across each state by week
week1AvgDeath <- sum(tracker$Avg.deaths[index1])
week2AvgDeath <- sum(tracker$Avg.deaths[index2])
week3AvgDeath <- sum(tracker$Avg.deaths[index3])
week4AvgDeath <- sum(tracker$Avg.deaths[index4])
week5AvgDeath <- sum(tracker$Avg.deaths[index5])
week6AvgDeath <- sum(tracker$Avg.deaths[index6])

# normalize across national pop
nationalPop <- 329389340

week1NatDeathRate <- week1AvgDeath/nationalPop
week2NatDeathRate <- week2AvgDeath/nationalPop
week3NatDeathRate <- week3AvgDeath/nationalPop
week4NatDeathRate <- week4AvgDeath/nationalPop
week5NatDeathRate <- week5AvgDeath/nationalPop
week6NatDeathRate <- week6AvgDeath/nationalPop

```

# calc rates for each state
```{r}
# calc death rate for each state
#setwd("C:/Users/ekw28/OneDrive - Yale University/Desktop/DM_COVID/data")
setwd("C:/Users/ekw28/ekw_lab_projects/dm_covid/dm_covid_current")
#state_pops <- read.xlsx("state_pops.xlsx")
#state_pops <- read_excel("C:/Users/ekw28/ekw_lab_projects/dm_covid/dm_covid_current/state_pops.xlsx")
state_pops <- read.csv("state_pop_2010_census.csv")
colnames(state_pops) <- c("state", "pop")

#state_pop <- state_pops2
state_pop <- as.data.frame(state_pops)
trackerNew <- merge(tracker,state_pops,by.x ="state") # trackerNew has cols for state pop & death rates

trackerNew$deathRate <- NA

for (i in 1:nrow(trackerNew)){
   trackerNew$deathRate[i] <- trackerNew$Avg.deaths[i]/trackerNew$pop[i]
}

# calc avg positive test rate for e/ state
trackerNew$posTestRate <- NA
for (i in 1:nrow(trackerNew)){
   trackerNew$posTestRate[i] <- trackerNew$Avg.positive.tests[i]/trackerNew$pop[i]
}

# calc avg hospitalization rate for e/ state
trackerNew$hospRate <- NA
for (i in 1:nrow(trackerNew)){
   trackerNew$hospRate[i] <- trackerNew$Avg.hospitalized[i]/trackerNew$pop[i]
}

```

# MODELING
# run GLM using death rate, testing, hospitalizations, age, and R&A task attitude 
```{r}
# put all data for GLM into 1 dataframe
modData <- data.frame(data.all$state, data.all$week, data.all$age, data.all$risk.mon, data.all$risk.med, data.all$ambig_corr.mon, data.all$ambig_corr.med, data.all$ambig1_react_state)

modData$death.rate <- NA
modData$hosp.rate <- NA
modData$pos.test.rate <- NA

# get external metrics from trackerNew for e/ sub in data.all

for (i in 1:nrow(modData)) {
  # find current state & week in modData
  currentState <- modData$data.all.state[i] # find current state
  currentWeek <- modData$data.all.week[i] # find current week

  # find matching state & week in trackerNew
  trackerNew_data <- filter(trackerNew, state == currentState & Week == currentWeek)
  
  # fill appropriate indices of modData with data from trackerNew_data
  modData$death.rate[i] <- trackerNew_data$deathRate
  modData$hosp.rate[i] <- trackerNew_data$hospRate
  modData$pos.test.rate[i] <- trackerNew_data$posTestRate
}

# GLM model 1: risk.mon ~ death rate + testing + hospitalizations + 
model1 <- glm(modData$data.all.risk.mon ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate) #+ modData$data.all.age)
summary(model1)
R# plot residuals
plot(density(resid(model1, type="response")))
sum(is.na(modData$hosp.rate))
ggplot(modData,aes(x=death.rate, y=hosp.rate)) + 
  geom_point()
plot(model1$fitted.values,resid(model1, type="response"))


# GLM model 2: ambig_corr.mon ~ death rate + testing + hospitalizations + 
model2 <- glm(modData$data.all.ambig_corr.mon ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate) #+ modData$data.all.age)
summary(model2)
# plot residuals
plot(model2$fitted.values,resid(model2, type="response"))

# GLM model 3: risk.med ~ death rate + testing + hospitalizations + 
model3 <- glm(modData$data.all.risk.med ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate) #+ modData$data.all.age)
# plot residuals
plot(density(resid(model3, type="response")))
plot(model3$fitted.values,resid(model3, type="response"))
summary(model3)

# model 4: ambig_corr.med ~ death rate + testing + hospitalizations + 
model4 <- glm(modData$data.all.ambig_corr.med ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate)# + modData$data.all.age)

# plot residuals
plot(density(resid(model4, type="response")))
plot(model4$fitted.values,resid(model4, type="response"))
summary(model4)


# GLM model 5: react state ~ hosp.rate + death.rate + pos.test.rate + ambib_corr.med + ambig_corr.mon + risk.mon + risk.med
model5GLM <- glm(modData$data.all.ambig1_react_state ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate + modData$data.all.risk.med + modData$data.all.risk.mon + modData$data.all.age)
summary(model5GLM)
# plot residuals
plot(model5GLM$fitted.values,resid(model5GLM, type="response"))
```

# add political typology, religiosity, current earnings, and household income to parameters in model 5

data.all$political_typology (1 = very liberal, 2 = liberal, 3 = slightly liberal, 4 = moderate, 5 = slightly conservative, 6 = conservative, 7 = very conservative, 8 = don't know/apolitical, 9 = libertarian, 10 = other), int

data.all$religiosity_rating (1 = very religious, 2 = moderately religious, 3 = slightly religious, 4 = not sure, 5 = slightly not religious, 6 = moderately not religious, 8 = not at all religious), int

data.all$household.earnings (what impact will pandemic have on household earnings in 2021? 1 = significantly decrease, 2 = moderately decrease, 3 = slightly decrease, 4 = no impact, 5 = slight increase, 6 = moderate increase, 7 = significant increase ), int

data.all$income_last.year (What was the total income of your household during the last year (2020), in your own currency? Fill in the blank), char


```{r}
# clean income_last.year -- only use numerics 
data.all$income_last.year <- as.character(data.all$income_last.year)
library(readr)
data.all$income_last.year <- parse_number(data.all$income_last.year)

# add new parameters into modData 
modData$political.typology <- data.all$political_typology
modData$religiosity <- data.all$religiosity_rating
modData$household.earnings <- data.all$household.earnings
modData$income.last.year <- data.all$income_last.year

# correct ordinal data --> numeric
modData$political.typology <- as.numeric(modData$political.typology)
modData$religiosity <- as.numeric(modData$religiosity)
modData$household.earnings <- as.numeric(modData$household.earnings)

# exclude political typology with ratings 8-10
modData$political.typology <- na_if(modData$political.typology, "8")
modData$political.typology <- na_if(modData$political.typology, "9")
modData$political.typology <- na_if(modData$political.typology, "10")





# GLM model 6: react state ~ hosp.rate + death.rate + pos.test.rate + ambib_corr.med + ambig_corr.mon + risk.mon + risk.med + religiosity + political typology + household earnings + income last year 

model6GLM <- glm(modData$data.all.ambig1_react_state ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate + modData$data.all.risk.med + modData$data.all.risk.mon + modData$data.all.age + modData$political.typology + modData$religiosity + modData$income.last.year + modData$household.earnings + modData$data.all.ambig_corr.med + modData$data.all.ambig_corr.mon)

summary(model6GLM)

# plot residuals
plot(model6GLM$fitted.values,resid(model6GLM, type="response"))

model6GLM <- glm(modData$data.all.ambig1_react_state ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate +  modData$data.all.age + modData$political.typology + modData$religiosity + modData$income.last.year + modData$household.earnings + modData$data.all.ambig_corr.med + modData$data.all.ambig_corr.mon)


```


```{r}
# GLM model 7: react_state ~ ambig med, ambig mon, risk med, risk mon
model7GLM <- glm(modData$data.all.ambig1_react_state ~ modData$data.all.risk.med + modData$data.all.risk.mon + modData$data.all.ambig_corr.med + modData$data.all.ambig_corr.mon)

summary(model7GLM)

```
# Large states only 

```{r}
large_states <- c("CA", "TX", "OH", "NY", "FL")
modData_large <- filter(modData, modData$data.all.state == large_states)
```


```{r}
# using only death rate as external metric    
model7GLM <- glm(modData$data.all.ambig1_react_state ~ modData$death.rate + modData$data.all.risk.med + modData$data.all.risk.mon + modData$data.all.age + modData$political.typology + modData$religiosity + modData$income.last.year + modData$data.all.ambig_corr.med + modData$data.all.ambig_corr.mon)

summary(model7GLM)

# using only hosp rate as external metric    
model8GLM <- glm(modData$data.all.ambig1_react_state ~ modData$hosp.rate + modData$data.all.risk.med + modData$data.all.risk.mon + modData$data.all.age + modData$political.typology + modData$religiosity + modData$income.last.year + modData$data.all.ambig_corr.med + modData$data.all.ambig_corr.mon)

summary(model8GLM)

# using only positive test rate
model9GLM <- glm(modData$data.all.ambig1_react_state ~ modData$pos.test.rate + modData$data.all.risk.med + modData$data.all.risk.mon + modData$data.all.age + modData$political.typology + modData$religiosity + modData$income.last.year + modData$data.all.ambig_corr.med + modData$data.all.ambig_corr.mon)

summary(model9GLM)
```

