---
title: "Covid Phase Two Analysis
output: html_notebook
---

# preprocessing

```{r}
rm(list = ls())

# Download package tarball from CRAN archive
url <- "https://cran.r-project.org/src/contrib/Archive/zipcode/zipcode_1.0.tar.gz"
pkgFile <- "zipcode_1.0.tar.gz"
download.file(url = url, destfile = pkgFile)
# Install dependencies
install.packages(c("ada", "ipred", "evd","sjmisc"))
# Install package
install.packages(pkgs=pkgFile, type="source", repos=NULL)
# Delete package tarball
unlink(pkgFile)

# import libraries
library(ggplot2)
library(zipcode) # install only available through archive
library(maps)
library(ggpubr)
library("PerformanceAnalytics") # for chart correlation
library(usmap) # for plotting data us maps
library(scales) # for muted() function
library(tidyverse)
library(dplyr)
# library(memisc)
library(MASS)
```

```{r}
# load data

#setwd("Y:/Lab_Projects/RA_Medical_covid19/behavioral/mdm_covid19_data")
setwd("C:/Users/ekw28/ekw_lab_projects/dm_covid/dm_covid_current")
data.all.2 <- read.csv("data_all_phase2_306sub_with_attitude.csv")
tracker <- read.csv("phase_2_state_weekly_avgs.csv")
#tracker <- all_state_avgs_2_clean
#state_name <- read.csv("state_name.csv")
state_name <- read.csv("state_name.csv")
state_pops <- read.csv("state_pop_2010_census.csv")
data.all.2$age <- as.numeric(as.character(data.all.2$age)) + 17 # correct coding
```


```{r}
# merge tracker and state name
# colnames(state_name)[2] <- "state"
# tracker <- merge(tracker, state_name, by="state")


# separate into old and young (cutoff = 60yo)
data.all.2$is.young <- 1
data.all.2$is.young[data.all.2$age >= 60] = 0 
data.all.2$is.young <- as.factor(data.all.2$is.young)

# clean zipcode
data.all.2$zip_postal_code <- as.numeric(as.character(data.all.2$zip_postal_code))
data(zipcode)
data.all.2$zipcode<- clean.zipcodes(data.all.2$zip_postal_code)

# determine state names using zip - i dont think this is working
colnames(zipcode)
colnames(zipcode)[1] = "zipcode"
data.all.2 <- merge(data.all.2, zipcode, by="zipcode")

```
# normalize avg deaths, hosps, and positive tests by state population
```{r}
colnames(state_pops) <- c("state", "pop")

trackerNew <- merge(tracker, state_pops, by.x ="state") # trackerNew has cols for state pop & death rates

trackerNew$deathRate <- NA

# calc death rate for e/ state
for (i in 1:nrow(trackerNew)){
   trackerNew$deathRate[i] <- trackerNew$avg.death[i]/trackerNew$pop[i]
}

# calc avg positive test rate for e/ state
trackerNew$posTestRate <- NA
for (i in 1:nrow(trackerNew)){
   trackerNew$posTestRate[i] <- trackerNew$avg.pos.test[i]/trackerNew$pop[i]
}

# calc avg hospitalization rate for e/ state
trackerNew$hospRate <- NA
for (i in 1:nrow(trackerNew)){
   trackerNew$hospRate[i] <- trackerNew$avg.hosp[i]/trackerNew$pop[i]
}
```

# sort data.all.2 into weeks corresponding w/ tracker 
```{r}
min(data.all.2$StartDate)
max(data.all.2$EndDate)

week <- c(1,2)
date <- c("2021-01-05", "2021-02-05")

timing <- data.frame(week,date)
timing$Week <- as.integer(timing$week)
tracker <- merge(tracker, timing, by="week")
```

# combine data.all.2 and tracker (look up how to do this more elegantly with lapply() )
```{r}
# change data type
tracker$state <- as.character(tracker$state)
tracker$week <- as.integer(as.character(tracker$week))

data.all.2$Avg.positive.tests <- NA
data.all.2$Avg.deaths <- NA
data.all.2$Avg.current.hospitalized <- NA
data.all.2$week <- NA

for (i in 1:nrow(data.all.2)) { 
  startDate <- as.numeric(format(as.Date(data.all.2$StartDate[i],format="%Y-%m-%d"), format = "%d"))
  data.all.2$week[i] <- ifelse(startDate > 20,1,2)
  
  state <- data.all.2$state[i]
  week <- data.all.2$week[i]
  data.all.2$avg.deaths[i] <- tracker$avg.death[tracker$state==state & tracker$week == week]
  data.all.2$Avg.current.hospitalized[i] <- tracker$avg.hosp[tracker$state==state & tracker$week == week]
  data.all.2$Avg.positive.tests[i] <- tracker$avg.pos.test[tracker$state==state & tracker$week == week]
}

```

# change data type
```{r}
data.all.2$ambig1_react_state <- as.integer(as.character(data.all.2$ambig1_react_state))
data.all.2$ambig1_gov_local <- as.integer(as.character(data.all.2$ambig1_gov_local))
data.all.2$anx_feelingnow <- as.integer(as.character(data.all.2$anx_feelingnow))
```

END PREPROCESSING

BEGIN ANALYSIS

# visualization
plot ambig1_react_ ratings v. age
```{r}
# ambig1_react_state v. age
react_state_ratings <- c("Completely overreacting", "Somewhat overreacting", "Appropriate", "Somewhat insufficient", "Completely insufficient") #1-5

data.all.2$ambig1_react_state <- as.factor(data.all.2$ambig1_react_state)

x1 <- ggplot(data.all.2, aes(x = ambig1_react_state, y = age)) + geom_boxplot() + ggtitle("ambig1_react_state v. age phase 2") + theme(plot.title = element_text(hjust = 0.5))

x1 + scale_x_discrete(labels = react_state_ratings) + theme(axis.text.x=element_text(angle=60,hjust=1))+ labs(x="rating", y="age")                                                                                                    
# ambig1_react_econ v. age                                                                                 
data.all.2$ambig1_react_econ <- as.factor(data.all.2$ambig1_react_econ)

x2 <- ggplot(data.all.2, aes(x = ambig1_react_econ, y = age)) + geom_boxplot() + 
  ggtitle("ambig1_react_econ v. age phase 2") + theme(plot.title = element_text(hjust = 0.5))

x2 + scale_x_discrete(labels = react_state_ratings) + theme(axis.text.x=element_text(angle=60,hjust=1))+ labs(x="rating", y="age")

plot(x2)

# ambig1_react_nation v. age                                                                              
data.all.2$ambig1_react_nation <- as.factor(data.all.2$ambig1_react_nation)

x3 <- ggplot(data.all.2, aes(x = ambig1_react_nation, y = age)) + geom_boxplot() + 
  ggtitle("ambig1_react_nation v. age phase 2") + theme(plot.title = element_text(hjust = 0.5))

x3 + scale_x_discrete(labels = react_state_ratings) + theme(axis.text.x=element_text(angle=60,hjust=1))+ labs(x="rating", y="age")

plot(x3)


# ambig1_react_science v. age                                                                          
data.all.2$ambig1_react_science <- as.factor(data.all.2$ambig1_react_science)

x4 <- ggplot(data.all.2, aes(x = ambig1_react_science, y = age)) + geom_boxplot() + 
  ggtitle("ambig1_react_science v. age phase 2") + theme(plot.title = element_text(hjust = 0.5))

x4 + scale_x_discrete(labels = react_state_ratings) + theme(axis.text.x=element_text(angle=60,hjust=1))+ labs(x="rating", y="age")

plot(x4)

# ambig1_react_worker v. age                                                                          
data.all.2$ambig1_react_worker <- as.factor(data.all.2$ambig1_react_worker)

x5 <- ggplot(data.all.2, aes(x = ambig1_react_worker, y = age)) + geom_boxplot() + 
  ggtitle("ambig1_react_worker v. age phase 2") + theme(plot.title = element_text(hjust = 0.5))

x5 + scale_x_discrete(labels = react_state_ratings) + theme(axis.text.x=element_text(angle=60,hjust=1))+ labs(x="rating", y="age")

plot(x5)

# ambig1_react_you v. age                                                                          
data.all.2$ambig1_react_you <- as.factor(data.all.2$ambig1_react_you)

x6 <- ggplot(data.all.2, aes(x = ambig1_react_you, y = age)) + geom_boxplot() + 
  ggtitle("ambig1_react_you v. age phase 2") + theme(plot.title = element_text(hjust = 0.5))

x6 + scale_x_discrete(labels = react_state_ratings) + theme(axis.text.x=element_text(angle=60,hjust=1))+ labs(x="rating", y="age")

plot(x6)

```
# plot ambig1_react_ v. political typology
```{r}
# ambig1_react_state v. politics
ambig1_react_ratings <- c("Completely overreacting", "Somewhat overreacting", "Appropriate", "Somewhat insufficient", "Completely insufficient") #1-5

political_typology_ratings <- c("Very liberal", "Liberal", "Slightly liberal", "Moderate", "Slightly conservative", "Conservative", "Very conservative", " Don't know", "Libertarian", "Other") #1-10

data.all.2$ambig1_react_state <- as.numeric(data.all.2$ambig1_react_state)
data.all.2$political_typology <- as.factor(data.all.2$political_typology)

x7 <- ggplot(data.all.2, aes(x = ambig1_react_state, y = political_typology)) + geom_violin() + ggtitle("ambig1_react_state v. political_typology phase 2") + theme(plot.title = element_text(hjust = 0.5))

x7 + stat_summary(fun = median, geom = "point") + scale_x_continuous(labels = react_state_ratings) + scale_y_discrete(labels = political_typology_ratings) + theme(axis.text.x=element_text(angle = 60,hjust = 1)) + labs(x = "react_state_rating", y = "political typology")

# ambig1_react_science v. pt
data.all.2$ambig1_react_science <- as.numeric(data.all.2$ambig1_react_science)

x8 <- ggplot(data.all.2, aes(x = ambig1_react_science, y = political_typology)) + geom_violin() + ggtitle("ambig1_react_science v. political_typology phase 2") + theme(plot.title = element_text(hjust = 0.5))

x8 + stat_summary(fun = median, geom = "point") + scale_x_continuous(labels = ambig1_react_ratings) + scale_y_discrete(labels = political_typology_ratings) + theme(axis.text.x=element_text(angle = 60,hjust = 1)) + labs(x = "react_science_rating", y = "political typology")

# ambig1_react_nation v. pt
data.all.2$ambig1_react_nation <- as.numeric(data.all.2$ambig1_react_nation)

x9 <- ggplot(data.all.2, aes(x = ambig1_react_nation, y = political_typology)) + geom_violin() + ggtitle("ambig1_react_nation v. political_typology phase 2") + theme(plot.title = element_text(hjust = 0.5))

x9 + stat_summary(fun = median, geom = "point") + scale_x_continuous(labels = ambig1_react_ratings) + scale_y_discrete(labels = political_typology_ratings) + theme(axis.text.x=element_text(angle = 60,hjust = 1)) + labs(x = "react_nation_rating", y = "political typology")

# ambig1_react_econ v. pt
data.all.2$ambig1_react_econ <- as.numeric(data.all.2$ambig1_react_econ)

x10 <- ggplot(data.all.2, aes(x = ambig1_react_econ, y = political_typology)) + geom_violin() + ggtitle("ambig1_react_econ v. political_typology phase 2") + theme(plot.title = element_text(hjust = 0.5))

x10 + stat_summary(fun = median, geom = "point") + scale_x_continuous(labels = ambig1_react_ratings) + scale_y_discrete(labels = political_typology_ratings) + theme(axis.text.x=element_text(angle = 60,hjust = 1)) + labs(x = "react_econ_rating", y = "political typology")

# ambig1_react_worker v. pt
data.all.2$ambig1_react_worker <- as.numeric(data.all.2$ambig1_react_worker)

x11 <- ggplot(data.all.2, aes(x = ambig1_react_worker, y = political_typology)) + geom_violin() + ggtitle("ambig1_react_worker v. political_typology phase 2") + theme(plot.title = element_text(hjust = 0.5))

x11 + stat_summary(fun = median, geom = "point") + scale_x_continuous(labels = ambig1_react_ratings) + scale_y_discrete(labels = political_typology_ratings) + theme(axis.text.x=element_text(angle = 60,hjust = 1)) + labs(x = "react_worker_rating", y = "political typology")

# ambig1_react_you v. pt
data.all.2$ambig1_react_you <- as.numeric(data.all.2$ambig1_react_you)

x12 <- ggplot(data.all.2, aes(x = ambig1_react_you, y = political_typology)) + geom_violin() + ggtitle("ambig1_react_you v. political_typology phase 2") + theme(plot.title = element_text(hjust = 0.5))

x12 + stat_summary(fun = median, geom = "point") + scale_x_continuous(labels = ambig1_react_ratings) + scale_y_discrete(labels = political_typology_ratings) + theme(axis.text.x=element_text(angle = 60,hjust = 1)) + labs(x = "react_you_rating", y = "political typology")
```
# MODELING
run GLM using death rate, testing, hospitalizations, age, and R&A task attitude 

```{r}
# put all data for GLM into 1 dataframe
modData <- data.frame(data.all.2$state, data.all.2$week, data.all.2$age, data.all.2$risk.mon, data.all.2$risk.med, data.all.2$ambig_corr.mon, data.all.2$ambig_corr.med, data.all.2$ambig1_react_state)

modData$death.rate <- NA
modData$hosp.rate <- NA
modData$pos.test.rate <- NA

# get external metrics from trackerNew for e/ sub in data.all.2
for (i in 1:nrow(modData)) {
  # find current state & week in modData
  currentState <- modData$data.all.2.state[i] # find current state
  currentWeek <- modData$data.all.2.week[i] # find current week

  # find matching state & week in trackerNew
  tracker_temp <- filter(trackerNew, state == currentState & week == currentWeek)
  
  # fill appropriate indices of modData with data from trackerNew_data
  modData$death.rate[i] <- tracker_temp$deathRate
  modData$hosp.rate[i] <- tracker_temp$hospRate
  modData$pos.test.rate[i] <- tracker_temp$posTestRate
}

```

task attitudes ~ external metrics
```{r}
# GLM model 1: risk.mon ~ death rate + testing + hospitalizations 
model1 <- glm(modData$data.all.2.risk.mon ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate )
summary(model1)

# GLM model 2: ambig_corr.mon ~ death rate + testing + hospitalizations
model2 <- glm(modData$data.all.2.ambig_corr.mon ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate)
summary(model2)

# GLM model 3: risk.med ~ death rate + testing + hospitalizations + age
model3 <- glm(modData$data.all.2.risk.med ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate)
summary(model3)

# model 4: ambig_corr.med ~ death rate + testing + hospitalizations 
model4 <- glm(modData$data.all.2.ambig_corr.med ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate)
summary(model4)
```
task attitudes ~ external metrics + age
```{r}
# GLM model 1: risk.mon ~ death rate + testing + hospitalizations 
model1 <- glm(modData$data.all.2.risk.mon ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate + modData$data.all.2.age)
summary(model1)

# GLM model 2: ambig_corr.mon ~ death rate + testing + hospitalizations
model2 <- glm(modData$data.all.2.ambig_corr.mon ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate + modData$data.all.2.age)
summary(model2)

# GLM model 3: risk.med ~ death rate + testing + hospitalizations + age
model3 <- glm(modData$data.all.2.risk.med ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate + modData$data.all.2.age)
summary(model3)

# model 4: ambig_corr.med ~ death rate + testing + hospitalizations 
model4 <- glm(modData$data.all.2.ambig_corr.med ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate + modData$data.all.2.age)
summary(model4)
```

ambig1_react ~ task attitude
```{r}
# GLM model 5: react state ~  ambig_corr.med + ambig_corr.mon + risk.mon + risk.med
model5GLM <- glm(modData$data.all.2.ambig1_react_state ~ modData$data.all.2.risk.med + modData$data.all.2.risk.mon + modData$data.all.2.ambig_corr.med + modData$data.all.2.ambig_corr.mon)
summary(model5GLM)
```
react state ~ task attitude + age
```{r}
# GLM model 5: react state ~  ambig_corr.med + ambig_corr.mon + risk.mon + risk.med
model5GLM <- glm(modData$data.all.2.ambig1_react_state ~ modData$data.all.2.risk.med + modData$data.all.2.risk.mon + modData$data.all.2.ambig_corr.med + modData$data.all.2.ambig_corr.mon + modData$data.all.2.age)
summary(model5GLM)
```
# add political typology, religiosity, current earnings, and household income into model 5

data.all$political_typology (1 = very liberal, 2 = liberal, 3 = slightly liberal, 4 = moderate, 5 = slightly conservative, 6 = conservative, 7 = very conservative, 8 = don't know/apolitical, 9 = libertarian, 10 = other), int

data.all$religiosity_rating (1 = very religious, 2 = moderately religious, 3 = slightly religious, 4 = not sure, 5 = slightly not religious, 6 = moderately not religious, 8 = not at all religious), int

data.all$household.earnings (what impact will pandemic have on household earnings in 2021? 1 = significantly decrease, 2 = moderately decrease, 3 = slightly decrease, 4 = no impact, 5 = slight increase, 6 = moderate increase, 7 = significant increase ), int

data.all$income_last.year (What was the total income of your household during the last year (2020), in your own currency? Fill in the blank), char

clean income
```{r}
# clean income_last.year -- only use numerics 
library(readr)

data.all.2$income_last.year <- as.character(data.all.2$income_last.year)
data.all.2$income_last.year <- parse_number(data.all.2$income_last.year)

# add new vars into modData 
modData$political.typology <- data.all.2$political_typology
modData$religiosity <- data.all.2$religiosity_rating
modData$household.earnings <- data.all.2$household.earnings
modData$income.last.year <- data.all.2$income_last.year

# correct ordinal data --> numeric
modData$political.typology <- as.numeric(modData$political.typology)
modData$religiosity <- as.numeric(modData$religiosity)
modData$household.earnings <- as.numeric(modData$household.earnings)

# exclude political typology with ratings 8-10
modData$political.typology <- na_if(modData$political.typology, "8")
modData$political.typology <- na_if(modData$political.typology, "9")
modData$political.typology <- na_if(modData$political.typology, "10")
```

react state ~ external metrics + task attitude + relg + $ + pt
```{r}
model6GLM <- glm(modData$data.all.2.ambig1_react_state ~ modData$death.rate + modData$pos.test.rate + modData$hosp.rate + modData$data.all.2.risk.med + modData$data.all.2.risk.mon + modData$data.all.2.age + modData$political.typology + modData$religiosity + modData$income.last.year + modData$household.earnings + modData$data.all.2.ambig_corr.med + modData$data.all.2.ambig_corr.mon)

summary(model6GLM)

```


# large states (incl. states as regressor)

```{r}
# put large states (n>= 20) from phase 1 into new dataframe
large_states_1 <- c("CA", "TX", "OH", "NY", "FL")

# find large states in phase 2
large_states_2 <- table(modData$data.all.2.state)
```

Large states in phase 1: NY, TX, OH, CA, FL
Large states in phase 2: NY, FL, OH, CA, PA 

  **what to do since large states aren't the same? ** --> use only states w/ n>20 & complete data: OH and FL

OH & FL w/ states as regressor phase 2
```{r}
# put into new df
#oh_fl_2 <- filter(modData, modData$data.all.2.state == "OH" & modData$data.all.2.state == "FL")
oh <- filter(modData, modData$data.all.2.state == "OH")
fl <- filter(modData, modData$data.all.2.state == "FL")
oh_fl_2 <- rbind(oh, fl)
```

debugging - why NAs in GLM summary? 
```{r}
# trying to see if regressors are perfectly correlated and that's why coefficients can't be estimated for e/ regressor
cor(oh_fl_2)
typeof(oh_fl_2)

# error 'x' must be numeric -- > need to use dummy vars for state# 0 = OH, 1 = FL
oh_fl_2$state_bin <- NA

for (i in 1:nrow(oh_fl_2)) { 
  state <- oh_fl_2$data.all.2.state[i]
  oh_fl_2$state_bin[i] <- ifelse(state == "OH",0,1) # 0 = OH, 1 = FL
}

#remove unecessary cols
oh_fl_2 <- subset(oh_fl_2, select = -c(data.all.2.state))
oh_fl_2 <- subset(oh_fl_2, select = -c(state_binary))

# remove NA from political typology
is.na(oh_fl_2)
na.omit(oh_fl_2)
is.na(oh_fl_2)

# retry cor to see if that's why NAs in model summary
install.packages("corrplot")
library(corrplot)
x <- cor(oh_fl_2, y = NULL, use = "complete.obs")
corrplot(x)

# must be a problem with death rate, pos test rate, hosp rate
```


repeat the large state glms from phase 1 on these two large states in phase 2
```{r}
# risk.mon ~ death rate + positive test rate + hosp rate + age + state
summary(glm(oh_fl_2$data.all.2.risk.mon ~ oh_fl_2$death.rate + oh_fl_2$pos.test.rate + oh_fl_2$hosp.rate + oh_fl_2$data.all.2.age + oh_fl_2$data.all.2.state))

# react_state ~ risk med + risk mon + ambig med + ambig mon + pos test + death + hosp

# react state ~ hosp.rate + death.rate + pos.test.rate + ambig_corr.med + ambig_corr.mon + risk.mon + risk.med + state

# react_state ~ death rate + positive test rate + hospitalization rate + risk med + risk mon + ambig med + ambig mon + age + political typology + religiosity + income last year + state 

# Mon ambig ~ death rate + positive test rate + hospitalization rate + age + state

# Med risk ~ death rate + positive test rate + hospitalization rate + age + state

# ambig_corr.med ~ death rate + positive test rate + hospitalization rate + age + state

# react state ~ hosp.rate + death.rate + pos.test.rate + ambig_corr.med + ambig_corr.mon + risk.mon + risk.med + age + state



```

  
trying w/ all 5 large states phase 2
```{r}
# create df for large states from e/ phase
modData_large_1 <- filter(modData, modData$data.all.2.state == large_states_1)

modData_large_2 <- subset(modData, data.all.2.state == large_states_2)
```

task attitudes ~ external metrics
  ** seems like death rate and pos test rate are perfectly correlated -> showing NA in glm output **
```{r}
# GLM model 1: risk.mon ~ death rate + testing + hospitalizations (phase 1)
summary(glm(modData_large_1$data.all.2.risk.mon ~ modData_large_1$death.rate + modData_large_1$pos.test.rate + modData_large_1$hosp.rate))

# GLM model 2: ambig_corr.mon ~ death rate + testing + hospitalizations
sumary((modData_large$data.all.2.ambig_corr.mon ~ modData_large$death.rate + modData_large$pos.test.rate + modData_large$hosp.rate))
# GLM model 3: risk.med ~ death rate + testing + hospitalizations + age
summary(glm(modData_large$data.all.2.risk.med ~ modData_large$death.rate + modData_large$pos.test.rate + modData_large$hosp.rate))

# model 4: ambig_corr.med ~ death rate + testing + hospitalizations 
summary(glm(modData_large$data.all.2.ambig_corr.med ~ modData_large$death.rate + modData_large$pos.test.rate + modData_large$hosp.rate))
```