---
title: "selected_questions_phase1_phase2_delta"
output: html_notebook
---

# import libraries
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
```



# data setup 
```{r}
setwd("Y:/Lab_Projects/RA_Medical_covid19/behavioral/mdm_covid19_data")

# import phase 1 & phase 2 data (R&A, survey, external metrics)
phase1 <- read.csv("phase1_RA_with_tracker.csv")
phase2 <- read.csv("phase2_RA_with_tracker.csv")

```
next: 
 - need to combine phase1 and phase 2 into 1 big df, but phase 1 and phase 2 are different dimensions. how to organize? 
    -need to determine which cols we want in the new, big df. 
      - should be the same # of cols for both after cleaning since we are only
        focusing on questions used in both phases
        
          include: 
          
            age
            week
            state
            
            ambig.mon
            ambig_corr.mon
            risk.med
            ambig.med                   
            ambig_corr.med
            
            Avg.positive.tests          
            Avg.total.test.results       
            Avg.deaths                   
            Avg.cumulative.hospitalized
            Avg.current.hospitalized    
            Avg.hospitalized.increase
            
            income last year
            politcal typology
            religiosity
            
            Are you currently covered by medical insurance?
            This week, about how much time per day have you spent consuming the news?
            How regularly have you been tracking the number of infections and deaths due to the COVID-19 pandemic?
            Washing hands
            Checking body temperature
            Dining out
            Attending religious services
            Meeting family and friends in person
            Using public transportation
            Stocking up on daily-life supplies
            Stocking up on prescription medication
            Are you wearing face masks or any other protective gear to cover your face, when outside of your house?
            Are you practicing social distancing?
            People may feel anxious under stressful situations, such as being on edge, nervous, worrying, easily annoyed, or having                trouble relaxing. How much anxiety in general are you feeling right now regarding the COVID-19 pandemic?
            How well-informed do you feel you are about the COVID-19 pandemic?
            
# combine into 1 df

does NOT include the questions that we identified as interesting but are only in phase 2 
  1.	How risk-taking do you think you are currently in daily activities (e.g. dining at restaurants, going to the gym, shopping at      stores, etc.), considering the potential exposure to viral infection?)	
  
  2. Compared with how you were in March 2020, how has the progression of the pandemic affected how risk-taking you are generally in     daily activities?		
  
  
```{r}
# select cols 

colnames_phase1 <- c('state', 'week', 'age', 'income_last.year', 'political_typology', 'religiosity_rating', 'ambig.mon', 'ambig_corr.mon', 'risk.med', 'ambig.med', 'ambig_corr.med', 'Avg.deaths', 'Avg.positive.tests', 'Average.current.hospitalized', 'medical.insurance', 'ambig1_news_perday', "ambig1_you", 'ambig1_track_deaths', 'ambig3_you_infect', 'ambig3_family_infect', 'actions_washinghands', 'actions_checkbodtemp', 'actions_diningout', 'actions)religoiusser', 'actions_fam_inperson', 'actions_usepubtrans', 'actions_stocsupplies', 'actions_stocmed', 'actions_wearmask', 'Q1089', 'anx_feelingnow') 

colnames_phase2 <- c('state', 'week', 'age', 'income_last.year', 'political_typology', 'religiosity_rating', 'ambig.mon', 'ambig_corr.mon', 'risk.med', 'ambig.med', 'ambig_corr.med', 'Avg.deaths', 'Avg.positive.tests', 'Average.current.hospitalized', 'medical.insurance', 'ambig1_news_perday', "ambig1_you", 'ambig1_track_deaths', 'ambig3_you_infect', 'ambig3_family_infect', 'actions_washinghands', 'actions_checkbodtemp', 'actions_diningout', 'actions)religoiusser', 'actions_fam_inperson', 'actions_usepubtrans', 'actions_stocsupplies', 'actions_stocmed', 'actions_wearmask', 'Q1089', 'anx_feelingnow') 

# new combined df
phase1_temp <- phase1[,colnames_phase1]
phase2_temp <- phase2[,colnames_phase2]

# cbind 

```
            
            
            
            
            

